---
title: Programming Assignment 3
author: <Umar Arshid - 301245546>
output:
 html_document:
 mathjax: default
---
```{r include=FALSE}

library(caTools)
require(caTools)
library("tree")
library("caret")
library(randomForest)
library("pROC")
install.packages('e1071', dependencies=TRUE)
```
## Task 1
```{r}
set.seed(1) 
train <- read.csv(file="Titanic.Train.csv",header=TRUE,sep=",");
test <- read.csv(file="Titanic.Test.csv",header=TRUE,sep=",");

```
 
##Task 1
```{r}

train$survived <- factor(train$survived)
test$survived <- factor(test$survived)
actual_results <- test$survived


frmla = survived ~ pclass + sex + age + parch + fare +sibsp
output.forest <- randomForest(frmla, data = train, importance=TRUE, ntree =100)
output_predict <- predict(output.forest, test, type = "class")
mean(actual_results == output_predict)
output_predict_int <- as.integer(as.character(output_predict)) 

rf.roc <- roc(actual_results, output_predict_int)
plot(rf.roc)
auc(rf.roc)
```

<report text for task 1 comes here>

Accuracy = 0.8396947
AUC = 0.8208

best accuracy in Assignment 2 is  0.6382978723 vs ~.80 from random forests

```{r}
## Task 2
importance(output.forest)
varImpPlot(output.forest)
```
<report text for task 2 comes here>
The Top 3 most important attributes are 
1. Sex
2. Age
3. Fare

Sex and age because women and children are always given priority in moments of evacuation and emergency

fare and pclass are related in the fact that the more you pay, generally the higher passenger class youll have. And the people who payed the most would be given higher priority in evacuation moments because 
A. higher the class the better the service around them
B. better evacuation methods.

```{r}
##Task 3

x.train <- model.matrix(frmla, data=train)
x.test <- model.matrix(frmla, data=test)
glmnFit <- train(x = x.train, 
             y = train$survived,
             method = "glmnet")
plot(glmnFit)
```
<report text for task 3 comes here>
pclass, parch, and fare are three most important attribute

```{r}
##Task 4
predictions <- predict(glmnFit, x.test)
confusionMatrix(predictions, actual_results)
output_prediction <- as.integer(as.character(predictions)) 
Lr.roc <- roc(actual_results, output_prediction)
plot(Lr.roc)
auc(Lr.roc)

```
<report text for task 4 comes here>
accuracy = 0.8092  
AUC - 0.7887

```{r}
library(e1071)
#Task 5
linear_tune <- tune(svm, survived ~., data=test, kernel="linear")
radial_tune <- tune(svm, survived ~., data=test, kernel="radial")
linear_tune$best.model
radial_tune$best.model

```
<report text for task 5 comes here>
Best parameters for linear svm is:
  SVM-Type:  C-classification 
  SVM-Kernel:  linear 
       cost:  1 
      gamma:  0.1428571 
BEst Parameters for radial svm is
Parameters:
   SVM-Type:  C-classification 
  SVM-Kernel:  radial 
       cost:  1 
      gamma:  0.1428571 
      
```{r}
  library(e1071)
  results.linear <- predict(linear_tune$best.model, test)
  results.radial <- predict(radial_tune$best.model, test)
  svmlin.roc <- roc(actual_results, as.numeric(results.linear))
  svmrad.roc <- roc(actual_results, as.numeric(results.radial))
  
  
  confusionMatrix(results.linear, actual_results)
  confusionMatrix(results.radial, actual_results)
  
  plot(svmlin.roc)
  auc(svmlin.roc)
  plot(svmrad.roc)
  auc(svmrad.roc)

  
```
<report text for task 6 comes here>
Linear acuracy = 0.8015 
Linear Area under the curve: 0.7842
Radial accuracy  = 0.8321 
Radial Area under the curve: 0.8145

